---
title: "3_Differential State Analysis"
output: html_notebook
---

#Differential State Analysis

#1-Set up 
```{r}
setwd("...")
getwd()

library(Seurat)
library(dplyr)
library(tidyr)
library(ggplot2)
library(muscat)
library(SingleCellExperiment)
```

#2-Loading total lung object 
```{r}
setwd("...")
seurat <- readRDS(file = "mouse_global_high_res.rds")

DimPlot(seurat, label = TRUE) + NoLegend()
```

```{r}
DefaultAssay(seurat) <- "RNA"
```


#3-Data Preparation

```{r}
seurat.sce <- as.SingleCellExperiment(seurat)

seurat.sce <- prepSCE(seurat.sce, 
                      kid = "CellType", #subpopulation assign
                      gid = "Sample", #treatment groups
                      sid = "Annotation", #biological reps
                      drop=T) #Drops all other colData columns

#Storing cluster and sample IDs + number of cluster and samples into simple variables for easy access
nk <- length(kids <- levels(seurat.sce$cluster_id))
ns <- length(sids <- levels(seurat.sce$sample_id))
names(kids) <- kids
names(sids) <- sids
```

#4-Data Overview

## Cluster Sample sizes

```{r}
# nb. of cells per cluster-sample
t(table(seurat.sce$cluster_id, seurat.sce$sample_id))
```

#5-Pseudo-bulk analysis

## Data Aggregation
```{r}
pb <- aggregateData(seurat.sce,
                    assay = "counts", fun = "sum",
                    by = c("cluster_id", "sample_id"))
# one sheet per subpopulation
assayNames(pb)
# pseudobulks for 1st subpopulation
t(head(assay(pb)))
```

Pseudo-bulk level MDS plot
```{r}
pb_mds <- pbMDS(pb)
pb_mds
```

```{r}
ei <- metadata(seurat.sce)$experiment_info
mm <- model.matrix(~0 + ei$group_id)
dimnames(mm) <- list(ei$sample_id, levels(ei$group_id))
contrast <- limma::makeContrasts("DPT_3d-Ctrl","DPT_5d-Ctrl","DTP_7d-Ctrl", levels=mm) # need to switch to the appropriate contrast Ctrl vs Treat_8wk or whatever
```

#6-Differential State Analysis

```{r}
res <- pbDS(pb, verbose = F, design = mm, contrast = contrast, min_cells = 1) #can add min.cells = 1, to include more groups in this analysis
```

#7-Filter and format results

```{r}
tbl_1 <- res$table[[1]] #generates a table of the results for the first contrast
tbl_2 <- res$table[[2]]
tbl_3 <- res$table[[3]]
tbl_4 <- res$table[[4]]
# one data.frame per cluster
names(tbl)
```

```{r}
# filter FDR < 5%, abs(logFC) > 1 & sort by adj. p-value
tbl_fil <- lapply(tbl, function(u) {
  u <- dplyr::filter(u, p_adj.loc < 0.05)
  dplyr::arrange(u, p_adj.loc)
})

# nb. of DS genes & % of total by cluster
n_de <- vapply(tbl_fil, nrow, numeric(1))
p_de <- format(n_de / nrow(seurat.sce) * 100, digits = 3)
data.frame("#DS" = n_de, "%DS" = p_de, check.names = FALSE)

# view top 2 hits in each cluster
top2 <- bind_rows(lapply(tbl_fil, top_n, 2, p_adj.loc))
format(top2[, -ncol(top2)], digits = 2)
```

```{r}
tmp <- seurat.sce
counts(tmp) <- as.matrix(counts(tmp))
result_table <- resDS(tmp, res, bind = "row", frq = FALSE, cpm = FALSE)
rm(tmp)
```

```{r}
count_mat <- as.matrix(seurat[["RNA"]]@data) > 0

cluster_list <- unique(result_table$cluster_id)

result_table$Ctrl.frq <- 0
result_table$DPT_3d.frq <- 0
result_table$DPT_5d.frq <- 0
result_table$DTP_7d.frq <- 0


Ctrl_cells <- colnames(seurat)[seurat$Sample == "Ctrl"]
for(i in 1:length(cluster_list)){
  #Which cells to calc freq on
  cluster_cells <- colnames(seurat)[seurat$CellType == cluster_list[i]]
  test_cells <- intersect(Ctrl_cells, cluster_cells)
  #Which rows of result_table to fill
  row_ind <- which(result_table$cluster_id == cluster_list[i])
  frq <- rowSums(count_mat[result_table$gene[row_ind],test_cells]) / length(test_cells)
  result_table$Ctrl.frq[row_ind] <- frq
}

DPT_3d_cells <- colnames(seurat)[seurat$Sample == "DPT_3d"]
for(i in 1:length(cluster_list)){
  #Which cells to calc freq on
  cluster_cells <- colnames(seurat)[seurat$CellType == cluster_list[i]]
  test_cells <- intersect(DPT_3d_cells, cluster_cells)
  #Which rows of result_table to fill
  row_ind <- which(result_table$cluster_id == cluster_list[i])
  frq <- rowSums(count_mat[result_table$gene[row_ind],test_cells]) / length(test_cells)
  result_table$DPT_3d.frq[row_ind] <- frq
}

DPT_5d_cells <- colnames(seurat)[seurat$Sample == "DPT_5d"]
for(i in 1:length(cluster_list)){
  #Which cells to calc freq on
  cluster_cells <- colnames(seurat)[seurat$CellType == cluster_list[i]]
  test_cells <- intersect(DPT_5d_cells, cluster_cells)
  #Which rows of result_table to fill
  row_ind <- which(result_table$cluster_id == cluster_list[i])
  frq <- rowSums(count_mat[result_table$gene[row_ind],test_cells]) / length(test_cells)
  result_table$DPT_5d.frq[row_ind] <- frq
}

DTP_7d_cells <- colnames(seurat)[seurat$Sample == "DTP_7d"]
for(i in 1:length(cluster_list)){
  #Which cells to calc freq on
  cluster_cells <- colnames(seurat)[seurat$CellType == cluster_list[i]]
  test_cells <- intersect(DTP_7d_cells, cluster_cells)
  #Which rows of result_table to fill
  row_ind <- which(result_table$cluster_id == cluster_list[i])
  frq <- rowSums(count_mat[result_table$gene[row_ind],test_cells]) / length(test_cells)
  result_table$DTP_7d.frq[row_ind] <- frq
}
```

## Save results table with the total EC DSA
```{r}
setwd("...")
write.csv(result_table, file = "mouse_dsa_results_global_high_res_min1cell.csv", row.names = F)
```

```{r}
filt_result_table <- result_table %>%
  filter(p_adj.loc <= 0.05) %>%
  filter(Ctrl.frq >= 0.05 | DPT_3d.frq >= 0.05 | DPT_5d.frq >= 0.05 | DTP_7d.frq >= 0.05)
```

```{r}
setwd("...")
write.csv(filt_result_table, file = "mouse_dsa_results_global_high_res_filtered_min1cell.csv", row.names = F)
```

#8-Visualization of DSA

```{r}
pbHeatmap(seurat.sce, res, top_n = 5)
```

```{r}
pbHeatmap(seurat.sce, res, k = "gCap-Trans.")
```

#9-GSEA - did not run this

```{r}
setwd("...")
result_table <- read.csv(file = "dsa_results_endothelial_all.csv")
```

```{r}
setwd("...")
hallmarks <- fgsea::gmtPathways("../data/hallmark.genesets.v6.1.symbols.gmt") #50 gene sets
kegg <- fgsea::gmtPathways("../data/kegg.genesets.v6.1.symbols.gmt") #186
go <- fgsea::gmtPathways("../data/GOTerms.BP.v6.1.symbols.gmt") #4436
reactome <- fgsea::gmtPathways("../data/reactome.genesets.v6.1.symbols.gmt") #674

gene_sets <- c(hallmarks, kegg, go, reactome)
```

```{r}
result_table1 <- result_table %>% filter(contrast == "Treat_1wk-Ctrl")
result_table3 <- result_table %>% filter(contrast == "Treat_3wk-Ctrl")
result_table5 <- result_table %>% filter(contrast == "Treat_5wk-Ctrl")
result_table8 <- result_table %>% filter(contrast == "Treat_8wk-Ctrl")
```

## 1wk vs Ctrl

```{r}
runGSEA <- function(cluster){
  results <- filter(result_table1, cluster_id == cluster)
  results <- filter(results, Ctrl.frq >= 0.05 |
                      Treat_1wk.frq >= 0.05)
  results <- arrange(results, desc(logFC))
  
  cluster_genes <- results$logFC
  names(cluster_genes) <- toupper(results$gene)
  
  gsea <- fgsea(pathways = gene_sets,
                stats = cluster_genes,
                minSize=15,
                maxSize=500,
                nproc = 2)
  gsea$cluster <- cluster
  
  return(gsea)
}
```

```{r}
cluster_list <- unique(result_table1$cluster_id)
fgsea_results <- lapply(cluster_list, runGSEA)
```

Format results
```{r}
fgsea_results <- do.call("rbind", fgsea_results)
fgsea_results <- as.data.frame(fgsea_results)
fgsea_results$leadingEdge <- as.character(fgsea_results$leadingEdge)
```


```{r}
writexl::write_xlsx(fgsea_results, path = "...")
```

Filter results for significant pathways
```{r}
gsea_sig <- filter(fgsea_results, padj <= 0.05) %>% arrange(NES)
```

```{r}
setwd("...")
write.csv(gsea_sig, file = "dsa_gsea_endothelial_1wk-Ctrl_results.csv")
```

```{r}
gsea_sig1wk <- gsea_sig
```

```{r}
ggplot(gsea_sig1wk, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=cluster)) +
  coord_flip() +
  labs(x="Pathway", y = "NES") +
  theme_minimal()
```

## 3wk vs Ctrl

```{r}
runGSEA <- function(cluster){
  results <- filter(result_table3, cluster_id == cluster)
  results <- filter(results, Ctrl.frq >= 0.05 |
                      Treat_3wk.frq >= 0.05)
  results <- arrange(results, desc(logFC))
  
  cluster_genes <- results$logFC
  names(cluster_genes) <- toupper(results$gene)
  
  gsea <- fgsea(pathways = gene_sets,
                stats = cluster_genes,
                minSize=15,
                maxSize=500,
                nproc = 2)
  gsea$cluster <- cluster
  
  return(gsea)
}
```

```{r}
cluster_list <- unique(result_table3$cluster_id)
fgsea_results <- lapply(cluster_list, runGSEA)
```

Format results
```{r}
fgsea_results <- do.call("rbind", fgsea_results)
fgsea_results <- as.data.frame(fgsea_results)
fgsea_results$leadingEdge <- as.character(fgsea_results$leadingEdge)
```

Filter results for significant pathways
```{r}
gsea_sig <- filter(fgsea_results, padj <= 0.05) %>% arrange(NES)
```

```{r}
setwd("...")
write.csv(gsea_sig, file = "dsa_gsea_endothelial_3wk-Ctrl_results.csv")
```

```{r}
gsea_sig3wk <- gsea_sig
```

## 5wk vs Ctrl

```{r}
runGSEA <- function(cluster){
  results <- filter(result_table5, cluster_id == cluster)
  results <- filter(results, Ctrl.frq >= 0.05 |
                      Treat_5wk.frq >= 0.05)
  results <- arrange(results, desc(logFC))
  
  cluster_genes <- results$logFC
  names(cluster_genes) <- toupper(results$gene)
  
  gsea <- fgsea(pathways = gene_sets,
                stats = cluster_genes,
                minSize=15,
                maxSize=500,
                nproc = 2)
  gsea$cluster <- cluster
  
  return(gsea)
}
```

```{r}
cluster_list <- unique(result_table5$cluster_id)
fgsea_results <- lapply(cluster_list, runGSEA)
```

Format results
```{r}
fgsea_results <- do.call("rbind", fgsea_results)
fgsea_results <- as.data.frame(fgsea_results)
fgsea_results$leadingEdge <- as.character(fgsea_results$leadingEdge)
```

Filter results for significant pathways
```{r}
gsea_sig <- filter(fgsea_results, padj <= 0.05) %>% arrange(NES)
```

```{r}
setwd("...")
write.csv(gsea_sig, file = "dsa_gsea_endothelial_5wk-Ctrl_results.csv")
```

```{r}
gsea_sig5wk <- gsea_sig
```

## 8wk vs Ctrl

```{r}
runGSEA <- function(cluster){
  results <- filter(result_table8, cluster_id == cluster)
  results <- filter(results, Ctrl.frq >= 0.05 |
                      Treat_8wk.frq >= 0.05)
  results <- arrange(results, desc(logFC))
  
  cluster_genes <- results$logFC
  names(cluster_genes) <- toupper(results$gene)
  
  gsea <- fgsea(pathways = gene_sets,
                stats = cluster_genes,
                minSize=15,
                maxSize=500,
                nproc = 2)
  gsea$cluster <- cluster
  
  return(gsea)
}
```

```{r}
cluster_list <- unique(result_table8$cluster_id)
fgsea_results <- lapply(cluster_list, runGSEA)
```

Format results
```{r}
fgsea_results <- do.call("rbind", fgsea_results)
fgsea_results <- as.data.frame(fgsea_results)
fgsea_results$leadingEdge <- as.character(fgsea_results$leadingEdge)
```

Filter results for significant pathways
```{r}
gsea_sig <- filter(fgsea_results, padj <= 0.05) %>% arrange(NES)
```

```{r}
setwd("...")
write.csv(gsea_sig, file = "dsa_gsea_endothelial_8wk-Ctrl_results.csv")
```

```{r}
gsea_sig8wk <- gsea_sig
```

#10-Visualize w Enrichment Plot

```{r}
BiocManager::install("enrichplot")
library(enrichplot)
```

## Load 1week comparison
```{r}
setwd("...")
gsea_sig1wk <- read.csv(file = "dsa_gsea_endothelial_1wk-Ctrl_results.csv")
```

```{r}
gsea_sig1wk <- mutate(gsea_sig1wk, log10_padj = -log10(padj))
```

```{r}
C0_1wk <- gsea_sig1wk %>% arrange(gsea_sig1wk, desc(NES)) %>% filter(cluster == "0")
top_10 <- top_n(C0_1wk, n=10, NES)
bot_10 <- top_n(C0_1wk, n=-10, NES)
C0_1wk <- full_join(top_10, bot_10)
```

```{r}
C_1wk <- gsea_sig1wk %>% arrange(gsea_sig1wk, desc(NES)) %>% filter(cluster == "1")
top_10 <- top_n(C_1wk, n=10, NES)
bot_10 <- top_n(C_1wk, n=-10, NES)
C1_1wk <- full_join(top_10, bot_10)
```

```{r}
C_1wk <- gsea_sig1wk %>% arrange(gsea_sig1wk, desc(NES)) %>% filter(cluster == "3")
top_10 <- top_n(C_1wk, n=10, NES)
bot_10 <- top_n(C_1wk, n=-10, NES)
C3_1wk <- full_join(top_10, bot_10)
```

```{r}
C_1wk <- gsea_sig1wk %>% arrange(gsea_sig1wk, desc(NES)) %>% filter(cluster == "4")
top_10 <- top_n(C_1wk, n=10, NES)
bot_10 <- top_n(C_1wk, n=-10, NES)
C4_1wk <- full_join(top_10, bot_10)
```

```{r}
C_1wk <- gsea_sig1wk %>% arrange(gsea_sig1wk, desc(NES)) %>% filter(cluster == "6")
top_10 <- top_n(C_1wk, n=10, NES)
bot_10 <- top_n(C_1wk, n=-10, NES)
C6_1wk <- full_join(top_10, bot_10)
```

```{r}
p1 <- ggplot(C0_1wk, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=log10_padj)) +
  coord_flip() +
  labs(x="Pathway", y = "NES") +
  theme_minimal()
p2 <- ggplot(C1_1wk, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=log10_padj)) +
  coord_flip() +
  labs(x="Pathway", y = "NES") +
  theme_minimal()
p3 <- ggplot(C3_1wk, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=log10_padj)) +
  coord_flip() +
  labs(x="Pathway", y = "NES") +
  theme_minimal()
p4 <- ggplot(C4_1wk, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=log10_padj)) +
  coord_flip() +
  labs(x="Pathway", y = "NES") +
  theme_minimal()
p5 <- ggplot(C6_1wk, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=log10_padj)) +
  coord_flip() +
  labs(x="Pathway", y = "NES") +
  theme_minimal()
```

View Plots
```{r}
p1
p2
p3
p4
p5
```

Save Plots
```{r}
setwd("...")
ggsave(p1, filename = "Cluster0_1wk_pathways.png", width=12, height= 7, dpi=300)
ggsave(p2, filename = "Cluster1_1wk_pathways.png", width=12, height= 7, dpi=300)
ggsave(p3, filename = "Cluster3_1wk_pathways.png", width=12, height= 7, dpi=300)
ggsave(p4, filename = "Cluster4_1wk_pathways.png", width=12, height= 7, dpi=300)
ggsave(p5, filename = "Cluster6_1wk_pathways.png", width=12, height= 7, dpi=300)
```


## Load 5week comparisons
```{r}
setwd("...")
gsea_sig5wk <- read.csv(file = "dsa_gsea_endothelial_5wk-Ctrl_results.csv")
```

```{r}
gsea_sig5wk <- mutate(gsea_sig5wk, log10_padj = -log10(padj))
```

```{r}
C <- gsea_sig5wk %>% arrange(gsea_sig5wk, desc(NES)) %>% filter(cluster == "0")
top_10 <- top_n(C, n=10, NES)
bot_10 <- top_n(C, n=-10, NES)
C0_5wk <- full_join(top_10, bot_10)
```

```{r}
C <- gsea_sig5wk %>% arrange(gsea_sig5wk, desc(NES)) %>% filter(cluster == "1")
top_10 <- top_n(C, n=10, NES)
bot_10 <- top_n(C, n=-10, NES)
C1_5wk <- full_join(top_10, bot_10)
```

```{r}
C <- gsea_sig5wk %>% arrange(gsea_sig5wk, desc(NES)) %>% filter(cluster == "3")
top_10 <- top_n(C, n=10, NES)
bot_10 <- top_n(C, n=-10, NES)
C3_5wk <- full_join(top_10, bot_10)
```

```{r}
C <- gsea_sig5wk %>% arrange(gsea_sig5wk, desc(NES)) %>% filter(cluster == "4")
top_10 <- top_n(C, n=10, NES)
bot_10 <- top_n(C, n=-10, NES)
C4_5wk <- full_join(top_10, bot_10)
```

```{r}
p1 <- ggplot(C0_5wk, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=log10_padj)) +
  coord_flip() +
  labs(x="Pathway", y = "NES") +
  theme_minimal()
p2 <- ggplot(C1_5wk, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=log10_padj)) +
  coord_flip() +
  labs(x="Pathway", y = "NES") +
  theme_minimal()
p3 <- ggplot(C3_5wk, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=log10_padj)) +
  coord_flip() +
  labs(x="Pathway", y = "NES") +
  theme_minimal()
p4 <- ggplot(C4_5wk, aes(reorder(pathway, NES), NES)) +
  geom_col(aes(fill=log10_padj)) +
  coord_flip() +
  labs(x="Pathway", y = "NES") +
  theme_minimal()
```

View Plots
```{r}
p1
p2
p3
p4
```

Save Plots
```{r}
setwd("...")
ggsave(p2, filename = "Cluster1_5wk_pathways.png", width=12, height= 7, dpi=300)
ggsave(p3, filename = "Cluster3_5wk_pathways.png", width=12, height= 7, dpi=300)
ggsave(p4, filename = "Cluster4_5wk_pathways.png", width=12, height= 7, dpi=300)
```